---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# activatr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/activatr)](https://cran.r-project.org/package=activatr)
[![R-CMD-check](https://github.com/dschafer/activatr/workflows/R-CMD-check/badge.svg)](https://github.com/dschafer/activatr/actions)
[![Codecov test coverage](https://codecov.io/gh/dschafer/activatr/branch/master/graph/badge.svg)](https://codecov.io/gh/dschafer/activatr?branch=master)
<!-- badges: end -->

`activatr` (pronounced like the word "activator") is a library for parsing GPX files into a standard format, and then manipulating and visualizing those files.

## Installation

You can install the released version of activatr from [CRAN](https://CRAN.R-project.org) with:

```{r install_release, eval = FALSE}
install.packages("activatr")
```

And the development version from [GitHub](https://github.com/) with:

```{r install_dev, eval = FALSE}
# install.packages("devtools")
devtools::install_github("dschafer/activatr")
```

## Example

Basic parsing of a GPX file is simple: we use the `parse_gpx` function and pass it the name of the GPX file.

```{r filename}
library(activatr)

# Get the running_example.gpx file included with this package.
filename <- system.file(
  "extdata",
  "running_example.gpx.gz",
  package = "activatr")

df <- parse_gpx(filename)
```

In its default configuration, `parse_gpx` will create a row for every GPS point in the file, and pull out the latitude (`lat`), longitude (`lon`), elevation (`ele`, in meters), and time (`time`) into the tibble:

```{r table, echo=FALSE, results='asis'}
knitr::kable(head(df, 5))
```

We can also get a summary of the activity:

```{r summary, results = "hide"}
summary(df)
```

```{r summary_table, echo=FALSE, results='asis'}
knitr::kable(summary(df))
```

Once we have the data, it's useful to visualize it. While basic visualizations work as expected with a data frame:

```{r basicplot, warning = FALSE}
library(ggplot2)
qplot(lon, lat, data=df)
```

It's more helpful to overlay this information on a correctly-sized map. To aid in that, `get_map_from_df` gives us a `ggmap` object (from the `ggmap` package), which we can use to visualize our track.

Let's see that on its own to start:

```{r mapplot_display, eval = FALSE}
ggmap::ggmap(get_ggmap_from_df(df)) + theme_void()
```

```{r mapplot_run, echo = FALSE}
# running_example_ggmap is the saved result of calling get_ggmap_from_df(df)
# We don't run that here because it requires an API key.
df_ggmap <- running_example_ggmap
ggmap::ggmap(running_example_ggmap) + theme_void()
```

The axes show that we now have a ggmap at the right size to visualize the run. So putting it
all together, we can make a nice basic graphic of the run:

```{r finalplot_display, eval = FALSE}
ggmap::ggmap(get_ggmap_from_df(df)) +
  theme_void() +
  geom_path(aes(x = lon, y = lat), size = 1, data = df, color = "red")
```

```{r finalplot_run, echo = FALSE}
ggmap::ggmap(running_example_ggmap) +
  theme_void() +
  geom_path(aes(x = lon, y = lat), size = 1, data = df, color = "red")
```
